Exploring the Link Between Family Background and Children's Academic Achievement in the United States (2021)

### 2023-04-05 outputs/paper/paper.qmd
intial draft only with few changes

---
title: "The Relationship between Family Status and Academic Achievement in Children in US from 2021"
author: 
  - Ruibo Sun
thanks: "https://github.com/ruibosun/how-parents-affect-childrens-education"
date: "`r Sys.time()`"
date-format: "D MMMM YYYY"
abstract: "This paper explores the relationship between family status variables, such as parents' socio-economic status and educational level, and their impact on children's academic achievement and education level using the data from 2021 General Social Survey (GSS). Previous studies have suggested that parents' education can positively influence their children's academic performance due to increased access to resources and greater involvement in their education. However, recent research has demonstrated that this relationship is complex and can be influenced by a variety of psychological and sociological factors. The findings suggest that parents' education, family income, and prestige score all have a positive correlation with children's education. Further research is needed to fully understand the complexities of this relationship and use the results to improve educational policies and programs."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

```{r,echo=FALSE}
#| include: false
#| warning: false
#| message: false
cleaned_data <- read_csv(
  file = here("inputs/data/GSS2021.csv"))
```

@@

@@

- **mapres10**: A numeric variable indicating the respondent's mother’s prestige score.
- **papres10**: A numeric variable indicating the respondent's father’s prestige score.
- **paeduc**: A numeric variable indicating the number of years of education completed by the respondent's father.
- **maeduc**: A numeric variable indicating the number of years of education completed by the respondent's mother.
- **educ**: A numeric variable indicating the number of years of education completed by the respondent.
- **mawrkslf**: A categorical variable indicating whether the respondent's mother was employed for wages or salary.
- **pawrkslf**: A categorical variable indicating whether the respondent's father was employed for wages or salary.
- **degree**: A categorical variable indicating the highest degree earned by the respondent.
- **born**: A categorical variable indicating the region of the US where the respondent was born.
- **sex**: A categorical variable indicating the respondent's sex.
- **reg16**: A categorical variable indicating the region of the US where the respondent lived at age of 16.
- **incom16**: A categorical variable indicating the respondent's total household at age of 16.

@@

# Results {#sec-result}

@@

# Model {#sec-model}

@@

# Fit the model
model <- lm(educ ~ mapres10 + papres10 + paeduc + maeduc, data = cleaned_data)

@@

# Discussion

@@

## Weaknesses and next steps {#sec-weakness}

>>>>

---
title: "Exploring the Link Between Family Background and Children's Academic Achievement in the United States (2021)"
author: 
  - Ruibo Sun
thanks: "Code and Data: https://github.com/ruibosun/revised-how-parents-affect-childrens-education"
date: "`r Sys.time()`"
date-format: "D MMMM YYYY"
abstract: "This paper explores the relationship between family status variables, such as parents' socio-economic status and educational level, and their impact on children's academic achievement and education level using the data from 2021 General Social Survey (GSS). Previous studies have suggested that parents' education can positively influence their children's academic performance due to increased access to resources and greater involvement in their education. However, recent research has demonstrated that this relationship is complex and can be influenced by a variety of psychological and sociological factors. The findings suggest that parents' education, family income, and prestige score all have a positive correlation with children's education. Further research is needed to fully understand the complexities of this relationship and use the results to improve educational policies and programs."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

```{r,echo=FALSE}
#| include: false
#| warning: false
#| message: false
cleaned_data <- read_csv(
  file = here("inputs/data/cleaned_GSS2021.csv"))
```

@@

In my previous paper, there are several findings. The average years of schooling in the US is 14, which is higher than the global average of 8.7 years. A bachelor's degree and a graduate degree are more common than an associate or junior college degree, and the US ranks behind several countries in terms of higher education attainment for 25-34 year-olds. Parental education levels have a positive correlation, and there is a strong link between parents' education levels and their children's academic achievement and societal status. Fathers' education levels appear to have a stronger impact than mothers' on children's academic achievement, but other factors may also play a role. Further research is needed to fully understand the relationship between these variables. Policymakers can use this information to make informed decisions about resource allocation and policy
development. Here is a [link](https://github.com/ruibosun/how-parents-affect-childrens-education/blob/main/outputs/paper/paper.pdf) to view the entire paper.However, the prior study did not appropriately handle the issue of missing values (NAs). Consequently, this new paper aims to conduct a more rigorous analysis by appropriately addressing the NAs. Moreover, it will address the limitations of the previous study to produce more accurate and representative results.

@@

- **mapres10**: A numeric variable indicating the respondent's mother’s prestige score.
- **papres10**: A numeric variable indicating the respondent's father’s prestige score.
- **paeduc**: A numeric variable indicating the number of years of education completed by the respondent's father.
- **maeduc**: A numeric variable indicating the number of years of education completed by the respondent's mother.
- **educ**: A numeric variable indicating the number of years of education completed by the respondent.
- **degree**: A categorical variable indicating the highest degree earned by the respondent.
- **born**: A categorical variable indicating the region of the US where the respondent was born.
- **sex**: A categorical variable indicating the respondent's sex.
- **reg16**: A categorical variable indicating the region of the US where the respondent lived at age of 16.
- **incom16**: A categorical variable indicating the respondent's total household at age of 16.

@@

# Results* {#sec-result}

@@

# Model* {#sec-model}

@@

# Fit the model
model <- lm(educ ~ mapres10 + papres10 + exp(paeduc) + exp(maeduc), data = cleaned_data)

@@

# Discussion*

@@

## Weaknesses and next steps* {#sec-weakness}
